---
title: "Exploratory analysis"
author: "Dereck de Mézquita"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), "./reports/", "/exploratory-analysis/exploratory-analysis.html")) })
output:
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
    # code_folding: hide
params:
  rmd: exploratory-analysis.Rmd
editor_options: 
  chunk_output_type: inline
---

<a download="exploratory-analysis.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 500, fig.width = 10, fig.height = 10, fig.align = "center")
```

# Libraries

```{r libraries, warning=FALSE, message=FALSE}
library("tidyverse")
options(dplyr.summarise.inform = FALSE)
# theme_set(theme(legend.position = "none"))
library("countrycode")
library("scales")
```

```{r modules, warning=FALSE, message=FALSE}
source("./modules/helper-functions.R")
```

# ggplot theme

```{r ggplot-theme}
theme_dereck <- function (base_size = 11, base_family = "") {
	theme_bw() %+replace% theme(
		plot.title = ggplot2::element_text(face = "bold", hjust = 1, margin = margin(0, 0, 5, 0, unit = "pt")), # 0.5 centres when position is set to "plot"
		plot.subtitle = ggplot2::element_text(hjust = 1, vjust = 0.015, margin = margin(0, 0, 5, 0, unit = "pt")),
		
		panel.background = ggplot2::element_blank(),
		# panel.border = ggplot2::element_rect(color = "#afafaf", fill = NA),
		axis.line = ggplot2::element_blank(),
		panel.border = ggplot2::element_blank(),
		
		panel.grid.minor = ggplot2::element_line(color = "#454545", size = 0.135, linetype = "dotted"),
		panel.grid.major = ggplot2::element_line(color = "#aaaaaa", size = 0.175, linetype = "longdash"),
		
		# axis.line = ggplot2::element_line(color = "#afafaf"),
		axis.ticks = ggplot2::element_line(color = "#aaaaaa"),
		
		strip.background = element_rect(fill = "white", colour = "#676767")
	)
}

theme_set(theme_dereck())
```

# Load data

```{r load-data}
data <- readRDS("./outputs/clean_data.Rds")
meta <- read.csv("./data/ddf--gapminder--fasttrack/meta-data/ddf--concepts.csv")
```

# Dataset completeness

```{r function-invert-na-df}
is.not.na <- function(x) {
	return(!is.na(x))
}
```

The datasets collected and cleaned are of the following dimensions (rows, columns):

1. `data$pop`: `r dim(data$pop)`
1. `data$countries`: `r dim(data$countries)`
1. `data$global`: `r dim(data$global)`
1. `data$world_4region`: `r dim(data$world_4region)`

## Tables preview {.tabset}

```{r preview-data-pop, results="asis"}
cat("### ","Dataset preview: pop", "\n")
datatable(rbind(head(data$manual, 50), tail(data$manual, 50)))
cat("\n\n")
```

```{r preview-data-countries, results="asis"}
cat("### ","Dataset preview: countries", "\n")
datatable(rbind(head(data$countries, 50), tail(data$countries, 50)))
cat("\n\n")
```

```{r preview-data-global, results="asis"}
cat("### ","Dataset preview: global", "\n")
datatable(rbind(head(data$global, 50), tail(data$global, 50)))
cat("\n\n")
```

```{r preview-data-world_4region, results="asis"}
cat("### ","Dataset preview: world_4region", "\n")
datatable(rbind(head(data$world_4region, 50), tail(data$world_4region, 50)))
cat("\n\n")
```

## Plots {.tabset}

```{r barplot-completion-pop, results="asis"}
cat("### ","Dataset completion: pop", "\n")
{data$manual %>%
		select(-Year, -Country, -Code, -region7, -region23, -eu15, -region7eu) %>%
		is.not.na() %>%
		as.data.frame() %>%
		colSums() / 58695 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	scale_x_discrete(expand = expansion(mult = c(0.015, 0.015))) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable") +
	coord_flip()
cat("\n\n")
```

```{r barplot-completion-countries, fig.height=15, results="asis"}
cat("### ","Dataset completion: countries", "\n")
{data$countries %>%
		select(-Country, -Year, -region7, -region7eu, -region23, -eu15) %>%
		is.not.na() %>%
		as.data.frame() %>%
		colSums() / 58369 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	scale_x_discrete(expand = expansion(mult = c(0.015, 0.015))) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable") +
	coord_flip()
cat("\n\n")
```

```{r barplot-completion-global, results="asis"}
cat("### ","Dataset completion: global", "\n")
{data$global %>%
		select(-Region, -Year) %>%
		is.not.na() %>% 
		as.data.frame() %>%
		colSums() / 124 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable") +
	coord_flip()
cat("\n\n")
```

```{r barplot-completion-world_4region, results="asis"}
cat("### ","Dataset completion: world_4region", "\n")
{data$world_4region %>%
		select(-Region, -Year) %>%
		is.not.na() %>%
		as.data.frame() %>%
		colSums() / 301 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable") +
	coord_flip()
cat("\n\n")
```


# manual

## pop_total

```{r pop-line-world}
data$manual %>%
	rename(Population = pop_total) %>%
	select(Year, Country, Code, region7eu, region23, Population) %>%
	ggplot(aes(Year, Population, group = Code, colour = Code)) +
	geom_line(alpha = 0.5, size = 0.65) +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "World population per country per region", subtitle = "Red dotted line: projected", x = "Year", y = "Population") +
	facet_wrap(~ region7eu, scales = "free") +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(n.breaks = 10, labels = function(x) {
		# paste0(substr(format(x, scientific = FALSE), 1, 2), " B")
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	theme(panel.grid.minor = element_blank())

data$manual %>%
	rename(Population = pop_total) %>%
	select(Year, Country, Code, region7eu, region23, Population) %>%
	ggplot(aes(Year, Population, group = Code, colour = Code)) +
	geom_line(alpha = 0.5, size = 0.65) +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "World population per country per region", subtitle = "Red dotted line: projected", x = "Year", y = "Population log2") +
	# lemon::facet_rep_wrap(~ region7, scales = "fixed", repeat.tick.labels = TRUE) +
	facet_wrap(~ region7eu, scales = "free") +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(trans = "log2", n.breaks = 10, labels = function(x) {
		# paste0(substr(format(x, scientific = FALSE), 1, 2), " B")
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	theme(panel.grid.minor = element_blank())

data$manual %>%
	rename(Population = pop_total) %>%
	select(Year, Country, Code, region7eu, region23, Population) %>%
	ggplot(aes(Year, Population, group = Code, colour = region7eu)) +
	geom_line(alpha = 0.65, size = 0.65) +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5)) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.75))) + # make alpha coloured legend darker
	labs(title = "World population per country; coloured per region", subtitle = "Red dotted line: projected\n", x = "Year", y = "Population log2") +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(trans = "log2", n.breaks = 10, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	})
```

```{r pop-line-per-regions-sums}
data$manual %>%
	rename(Population = pop_total) %>%
	group_by(Year, region7eu) %>%
	summarise(region_sum = sum(Population)) %>%
	ggplot(aes(Year, region_sum, colour = region7eu)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	directlabels::geom_dl(aes(label = region7eu), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = region7eu), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 17, expand = expansion(mult = c(0.275, 0.275))) +
	scale_y_continuous(n.breaks = 17, labels = function(x) {
		# paste0(substr(format(x, scientific = FALSE), 1, 2), " B")
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	labs(title = "World population per region", subtitle = "Red dotted line: projected", x = "Year", y = "Population")

data$manual %>%
	rename(Population = pop_total) %>%
	group_by(Year, region23) %>%
	summarise(region_sum = sum(Population)) %>%
	ggplot(aes(Year, region_sum, colour = region23)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	directlabels::geom_dl(aes(label = region23), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = region23), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 17, expand = expansion(mult = c(0.275, 0.275))) +
	scale_y_continuous(n.breaks = 17, labels = function(x) {
		# paste0(substr(format(x, scientific = FALSE), 1, 2), " B")
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	labs(title = "World population per region", subtitle = "Red dotted line: projected", x = "Year", y = "Population")
```

```{r pop-line-per-total-sums-int}
data$manual %>%
	rename(Population = pop_total) %>%
	group_by(Year) %>%
	summarise(Population = sum(Population)) %>%
	mutate(colour = ifelse(Year <= 2020, "#444444", "red"),
			 type = ifelse(Year <= 2020, "solid", "dashed")) %>%
	ggplot(aes(Year, Population, colour = colour, linetype = type)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_color_identity() +
	scale_linetype_identity() +
	scale_x_continuous(n.breaks = 25) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	labs(title = "World population", subtitle = "Red dotted line: projected", x = "Year", y = "Population (Billions)") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank())

data$manual %>%
	rename(Population = pop_total) %>%
	group_by(Year) %>%
	summarise(Population = sum(Population)) %>%
	mutate(colour = ifelse(Year <= 2020, "#444444", "red"),
			 type = ifelse(Year <= 2020, "solid", "dashed")) %>%
	ggplot(aes(Year, log2(Population), colour = colour, linetype = type)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_color_identity() +
	scale_linetype_identity() +
	scale_x_continuous(n.breaks = 25) +
	scale_y_continuous(n.breaks = 25, labels = function(x) { # calculate log2 scale labels manually; set aes(log2(data))
		scales::label_number_si(accuracy = 0.1)(2 ^ x) # more n.breaks when using log2 scale
	}) +
	labs(title = "World population", subtitle = "Red dotted line: projected", x = "Year", y = "Population (Billions) log2") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank())
```

# gdp

PPP GDP is gross domestic product converted to international dollars using purchasing power parity rates. A PPP dollar has the same purchasing power over GDP as the U.S. dollar has in the United States. Data are in constant 2005 international dollars.

```{r}
data$manual %>%
	rename(gdp = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(gdp)) %>%
	ggplot(aes(Year, gdp, group = Code, colour = region7eu)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 25, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 20, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	labs(title = "GDP per country", subtitle = "Coloured by region\nConstant 2005 international dollars", x = "Year", y = "Total GDP") +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95)))

data$manual %>%
	rename(gdp = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(gdp)) %>%
	ggplot(aes(Year, log2(gdp), group = Code, colour = region7eu)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 30, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	labs(title = "GDP per country", subtitle = "Coloured by region\nConstant 2005 international dollars", x = "Year", y = "Total GDP log2") +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95)))
```

```{r dpi=750}
data$manual %>%
	rename(gdp = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(gdp)) %>%
	ggplot(aes(Year, gdp, group = Code, colour = Code)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 12, expand = expansion(mult = c(0.2, 0.2))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	labs(title = "GDP per country", subtitle = "Coloured by region\nConstant 2005 international dollars", x = "Year", y = "Total GDP") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(panel.grid.minor = element_blank()) +
	facet_wrap(~ region7eu, scales = "free")

data$manual %>%
	rename(gdp = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(gdp)) %>%
	ggplot(aes(Year, log2(gdp), group = Code, colour = Code)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.2, 0.2))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	labs(title = "GDP per country", subtitle = "Coloured by region\nConstant 2005 international dollars", x = "Year", y = "Total GDP log2") +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(panel.grid.minor = element_blank()) +
	facet_wrap(~ region7eu, scales = "free")
```

```{r}
data$manual %>%
	rename(gdp_cap = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(gdp_cap)) %>%
	ggplot(aes(Year, gdp_cap, group = Code, colour = region7eu)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 30, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	labs(title = "GDP/income income per capita per country", subtitle = "Coloured by region\nConstant 2011 international dollars", x = "Year", y = "GDP/income per capita") +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95)))

data$manual %>%
	rename(gdp_cap = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(gdp_cap)) %>%
	ggplot(aes(Year, log2(gdp_cap), group = Code, colour = region7eu)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 30, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	labs(title = "GDP/income per capita per country", subtitle = "Coloured by region\nConstant 2011 international dollars", x = "Year", y = "GDP/income per capita log2") +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95)))
```


```{r}
data$manual %>%
	rename(gdp_cap = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(gdp_cap)) %>%
	ggplot(aes(Year, gdp_cap, group = Code, colour = Code)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.2, 0.2))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	labs(title = "GDP/income per capita per region", subtitle = "Coloured by region\nConstant 2011 international dollars", x = "Year", y = "GDP/income per capita") +
	theme(panel.grid.minor = element_blank(), legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	facet_wrap(~ region7eu, scales = "free")

data$manual %>%
	rename(gdp_cap = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(gdp_cap)) %>%
	ggplot(aes(Year, log2(gdp_cap), group = Code, colour = Code)) +
	geom_line(alpha = 0.65, size = 0.65) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.2, 0.2))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	labs(title = "GDP/income per capita per region", subtitle = "Coloured by region\nConstant 2011 international dollars", x = "Year", y = "GDP/income per capita log2") +
	theme(panel.grid.minor = element_blank(), legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	scale_colour_discrete(name = "World region") +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	facet_wrap(~ region7eu, scales = "free")
```


## avg_life_expect_yrs

```{r}
data$manual %>%
	rename(Value = avg_life_expect_yrs) %>%
	select(Year, Code, region7eu, region23, Value) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = Code)) +
	geom_line(alpha = 0.65, size = 0.65) +
	geom_vline(xintercept = 2020, colour = "red", linetype = "dashed") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.2, 0.2))) +
	scale_y_continuous(n.breaks = 15) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	guides(colour =guide_legend(override.aes = list(alpha = 0.95))) +
	labs(title = "Average life expectancy per country", subtitle = "Coloured by region", x = "Year", y = "Average life expectancy") +
	facet_wrap(~ region7eu, scales = "free")
```



# Exploratory analysis

Here is the meta data associated to these variables.

```{r table-meta-preview}
largeTable(meta, targs = c(3, 7, 9, 12))
```

# world_4region

## wn_bothhouses_c

Percentage of parliamentary seats held by women.

wn_bothhouses_c	measure	http://gapm.io/dwparl	v2	December 15 2020	Pecentage of women in parliaments	Pecentage of women in parliaments	Pecentage of women in parliaments	Percentage of national parliamentary seats held by women. Lower and upper houses combined	percent	governance

```{r w4-women-held-line, fig.width=10, fig.height=10}
data$world_4region %>%
	select(Region, Year, wn_bothhouses_c) %>%
	rename(Value = wn_bothhouses_c) %>%
	ggplot(aes(Year, Value, colour = Region)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(n.breaks = 10, labels = function(x) paste0(x, "%")) + # Add percent sign
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Percentage of parliamentary seats held by women", x = "Year", y = "Women held parliamentary seats %")
```

## refugee_in_perc_wb

Share refugees	Refugees as share of population in the country of residence.

refugee_in_perc_wb	measure	http://gapm.io/drefugee_wb	v2	July 27 2020	Share refugees	Refugees as share of population in the country of residence.	Share refugees	This is the share of refugees as a percentage of total population in the country of residence.	percent	refugees

```{r w4-refugee-line, fig.width=10, fig.height=10,}
data$world_4region %>%
	select(Region, Year, refugee_in_perc_wb) %>%
	rename(Value = refugee_in_perc_wb) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Region)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(n.breaks = 10, labels = function(x) paste0(x, "%")) + # Add percent sign
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Percentage of refugees to total population of origin", x = "Year", y = "Refugees to total population of origin %") +
	scale_colour_discrete(name = "World region") # Customise legend title and variable
```

# ihme_hiv_death

This is the proportion of deaths caused by HIV/AIDS

ihme_hiv_death	measure	http://gapm.io/dcauses_death_ihme	v1	June 15 2020	HIV/AIDS deaths (%)	HIV/AIDS deaths (%)	HIV/AIDS deaths (%)	This is the proportion of deaths caused by HIV/AIDS (IHME)	percent	health

```{r w4-hiv-deaths-line}
data$world_4region %>%
	select(Region, Year, ihme_hiv_death) %>%
	rename(Value = ihme_hiv_death) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Region)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(n.breaks = 10, labels = function(x) paste0(x, "%")) + # Add percent sign
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Proportion of HIV/AIDS deaths", x = "Year", y = "Proportion of HIV/AIDS deaths %") +
	scale_colour_discrete(name = "World region")
```

```{r w4-hiv-deaths-log-line, fig.width=10, fig.height=10}
data$world_4region %>%
	select(Region, Year, ihme_hiv_death) %>%
	rename(Value = ihme_hiv_death) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Region)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(trans = "log2", n.breaks = 10, labels = function(x) paste0(x, "%")) + # Add percent sign
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Proportion of HIV/AIDS deaths", x = "Year", y = "Proportion of HIV/AIDS deaths %") +
	scale_colour_discrete(name = "World region")
```


## ihme_conflict_death

This is the proportion of deaths caused by conflict and terrorism

ihme_conflict_death	measure	http://gapm.io/dcauses_death_ihme	v1	June 15 2020	Conflict and terrorism deaths (%)	Conflict and terrorism deaths (%)	Conflict and terrorism deaths (%)	This is the proportion of deaths caused by conflict and terrorism (IHME)	percent	health


```{r w4-conflict-death-line}
data$world_4region %>%
	select(Region, Year, ihme_conflict_death) %>%
	rename(Value = ihme_conflict_death) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Region)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(n.breaks = 10, labels = function(x) paste0(x, "%")) + # Add percent sign
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Proportion of conflict/terrorism deaths", x = "Year", y = "Proportion of conflict/terrorism deaths %") +
	scale_colour_discrete(name = "World region")
```


```{r w4-conflict-death-log-line}
data$world_4region %>%
	select(Region, Year, ihme_conflict_death) %>%
	rename(Value = ihme_conflict_death) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Region)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(trans = "log2", n.breaks = 15, labels = function(x) paste0(round(x, digits = 5), "%")) + # Add percent sign
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Proportion of conflict/terrorism deaths", x = "Year", y = "Proportion of conflict/terrorism deaths % (log2)") +
	scale_colour_discrete(name = "World region")
```


## ihme_cardio_death

This is the proportion of deaths caused by Cardiovascular diseases

ihme_cardio_death	measure	http://gapm.io/dcauses_death_ihme	v1	June 15 2020	Cardiovascular diseases deaths (%)	Cardiovascular diseases deaths (%)	Cardiovascular diseases deaths (%)	This is the proportion of deaths caused by Cardiovascular diseases (IHME)	percent	health

```{r w4-cardio-deaths-line}
data$world_4region %>%
	select(Region, Year, ihme_cardio_death) %>%
	rename(Value = ihme_cardio_death) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Region)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Region), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) +
	scale_y_continuous(n.breaks = 10, labels = function(x) paste0(x, "%")) + # Add percent sign
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Proportion of cardiovascular disease deaths", x = "Year", y = "Proportion of cardiovascular disease deaths %") +
	scale_colour_discrete(name = "World region")
```

# global

## ndisasd_tot

Natural disaster deaths	Natural disaster deaths per year

ndisasd_tot	measure	http://gapm.io/dnatdisasters	v3	March 24 2020	Natural disaster deaths	Natural disaster deaths per year	Natural disaster deaths	This is the number of deaths from all natural disasters. The data includes all categories classified as "natural disasters" (distinguished from technological disasters, such as transport, industrial accidents,etc.. and complex disasters which include some major famine situation for which the drought were not the main causal factor). This includes total deaths from animal accident, droughts, earthquake, epidemic, extreme temperature, Flood, Fog, Impact, Insect infestation, Landslide, Mass movement (dry), Storm, Volcanic activity and Wildfire.	deaths	environment

```{r global-ndisaster-tot-bar}
data$global %>%
	select(Year, Region, ndisasd_tot, ndisasd_rate) %>%
	rename(Value = ndisasd_tot) %>%
	filter(!is.na(Value)) %>%
	filter(Year <= 2020) %>%
	ggplot(aes(Year, Value)) + # , fill = ndisasd_rate
	geom_bar(stat = "identity") +
	scale_x_continuous(n.breaks = 20) +
	scale_y_continuous(n.breaks = 15) +
	labs(title = "Total deaths related to natural disasters", x = "Year", y = "Total deaths related to natural disaster") +
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

## ndisasd_rate

This is the number of deaths from all natural disasters per million people	deaths	environment

ndisasd_rate	measure	http://gapm.io/dnatdisasters	v3	March 24 2020	Natural disasters death rate	Natural disasters death rate per million people per year	Natural disasters death rate	This is the number of deaths from all natural disasters per million people	deaths	environment

```{r global-ndisaster-rate-bar}
data$global %>%
	select(Year, Region, ndisasd_tot, ndisasd_rate) %>%
	rename(Value = ndisasd_rate) %>%
	filter(!is.na(Value)) %>%
	filter(Year <= 2020) %>%
	ggplot(aes(Year, (Value / 1000000) * 100)) + # , fill = ndisasd_rate
	geom_bar(stat = "identity") +
	scale_x_continuous(n.breaks = 25) +
	scale_y_continuous(n.breaks = 15, labels = function(x) paste0(x, "%")) +
	labs(title = "Proportion of natural disaster associated deaths to world population", x = "Year", y = "Proportion of natural disaster deaths to world population") +
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

## num_countries_fhos

How many countries have had a female leader

num_countries_fhos	measure	http://gapm.io/dfemale_hos_over_time	v1	May 3 2020	Female Head of State	Number of countries that have had Female Head of State	Female Head of State	This answer the question: How many countries have had a female leader? This list includes women who have been elected or appointed head of state or government of their respective countries since the interwar period. The list does not include women chosen by a hereditary monarch nor females who have been Prime Minister but not head of state or government (like Peru where only President is head of Government).	countries	governance


```{r global-female-head-line}
data$global %>%
	select(Year, num_countries_fhos) %>%
	rename(Value = num_countries_fhos) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value)) +
	geom_line() +
	scale_x_continuous(n.breaks = 25) +
	scale_y_continuous(n.breaks = 15) +
	labs(title = "Number of countries who have had a female head of state", x = "Year", y = "Countries that had a female head of state") +
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

## journakilled

journakilled	measure	http://gapm.io/djournalists_silenced	v1	January 22 2019	Journalists killed	Number of journalists killed

```{r global-journalist-killed-line}
data$global %>%
	select(Year, journakilled) %>%
	rename(Value = journakilled) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value)) +
	geom_line() +
	scale_x_continuous(n.breaks = 25) +
	scale_y_continuous(n.breaks = 20) +
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Total number of journalist murdered internationally", x = "Year", y = "Number of murdered journalist")
```

# countries

## gini

Gini shows income inequality in a society. A higher number means more inequality.

gini	measure	http://gapm.io/ddgini	v3	December 5 2018	Gini	Gini coefficient		Gini shows income inequality in a society. A higher number means more inequality.

```{r countries-inequality-line}
data$countries %>%
	select(region7eu, Year, gini) %>%
	rename(Value = gini) %>%
	filter(!is.na(Value) & Year <= 2020, !is.na(region7eu)) %>%
	group_by(Year, region7eu) %>%
	summarise(Value = mean(Value), region7eu) %>%
	ggplot(aes(Year, Value, group = region7eu, colour = region7eu)) +
	geom_line() +
	# geom_vline(xintercept = 2020, linetype = "dashed", colour = "red") +
	directlabels::geom_dl(aes(label = region7eu), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = region7eu), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Average gini per region (inequality index)", subtitle = "A higher number means more inequality", x = "Year", y = "Gini index value") +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.375, 0.375))) +
	scale_y_continuous(n.breaks = 25)
	# facet_wrap(~ Region)
```


```{r countries-inequality-facet-line}
data$countries %>%
	select(region7eu, Code, Country, Year, gini) %>%
	rename(Value = gini) %>%
	filter(!is.na(Value)) %>%
	group_by(Year, region7eu, Code) %>%
	summarise(Value = mean(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	labs(title = "Average gini per region (inequality index)", subtitle = "A higher number means more inequality", x = "Year", y = "Gini index value") +
	facet_wrap(~ region7eu, scales = "free") +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.375, 0.375))) +
	scale_y_continuous(n.breaks = 25)
```


<!-- ```{r fig.width=20, fig.height=20} -->
<!-- data$countries %>% -->
<!-- 	select(Region, Code, Country, Year, gini) %>% -->
<!-- 	rename(Value = gini) %>% -->
<!-- 	filter(!is.na(Value)) %>% -->
<!-- 	ggplot(aes(Year, Value, group = Code, colour = Code)) + -->
<!-- 	geom_line() + -->
<!-- 	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) + -->
<!-- 	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) + -->
<!-- 	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + -->
<!-- 	labs(title = "Gini (inequality index)", subtitle = "A higher number means more inequality", x = "Year", y = "Gini index value") + -->
<!-- 	facet_wrap(~ Region, scales = "free") + -->
<!-- 	# theme(axis.line = element_line()) + -->
<!-- 	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) + -->
<!-- 	scale_y_continuous(limits = c(15, 85), n.breaks = 25) -->

<!-- data$countries %>% -->
<!-- 	select(Region, Code, Country, Year, gini) %>% -->
<!-- 	rename(Value = gini) %>% -->
<!-- 	filter(!is.na(Value)) %>% -->
<!-- 	ggplot(aes(Year, Value, group = Code, colour = Code)) + -->
<!-- 	geom_line() + -->
<!-- 	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) + -->
<!-- 	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) + -->
<!-- 	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + -->
<!-- 	labs(title = "Gini (inequality index)", subtitle = "A higher number means more inequality", x = "Year", y = "Gini index value") + -->
<!-- 	facet_wrap(~ Region, scales = "free") + -->
<!-- 	# theme(axis.line = element_line()) + -->
<!-- 	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.15, 0.15))) + -->
<!-- 	scale_y_continuous(trans = "log10", n.breaks = 25) -->
<!-- ``` -->


## u5pop

Total number of children, aged 0 to 4 years old

u5pop	measure	http://gapm.io/du5pop	v1	January 31 2020	Under-five population	Number of under five years children	Under-five population	Total number of children, aged 0 to 4 years old.	people	population

```{r countries-u5pop-line}
data$countries %>%
	select(region7eu, Code, Year, u5pop) %>%
	rename(Value = u5pop) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	scale_x_continuous(n.breaks = 20,  expand = expansion(mult = c(0.2, 0.2))) +
	facet_wrap(~ region7eu, scales = "free") +
	labs(title = "Total number of children 0 - 5 yrs", x = "Year", y = "Number")

data$countries %>%
	select(region7eu, Code, Year, u5pop) %>%
	rename(Value = u5pop) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	scale_x_continuous(n.breaks = 20,  expand = expansion(mult = c(0.2, 0.2))) +
	facet_wrap(~ region7eu, scales = "free") +
	labs(title = "Total number of children 0 - 5 yrs", x = "Year", y = "Number")
```

### u5pop/population

```{r countries-u5pop-prc}
data$countries %>% 
	select(region7eu, Code, Year, u5pop, pop_total) %>%
	filter(!is.na(u5pop) & !is.na(pop_total)) %>%
	mutate(Value = ((u5pop / pop_total) * 100)) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7eu)) +
	geom_line(alpha = 0.35, size = 0.85) +
	geom_vline(xintercept = 2020, colour = "red", linetype = "dashed") +
	scale_x_continuous(n.breaks = 20) +
	scale_y_continuous(n.breaks = 10, labels = function(x) paste0(x, "%")) + # Add percent sign
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	guides(color = guide_legend(override.aes = list(alpha = 0.75))) + # make alpha coloured legend darker
	labs(title = "Percentage to country's population of children 0 - 5 yrs", x = "Year", y = "Percent of children under 5 yrs")
```

## female_hos

```{r countries-1-fhos}
data$countries %>%
	select(region7eu, Code, Year, female_hos) %>%
	filter(!is.na(female_hos) & Year <= 2019, !is.na(region7eu)) %>%
	mutate(Value = ifelse(female_hos == "Never had a female head of state", 0, 1)) %>%
	ggplot(aes(Year, Value, group = region7eu, fill = region7eu)) +
	geom_bar(stat = "identity") +
	scale_y_continuous(n.breaks = 15) +
	scale_x_continuous(n.breaks = 20) +
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Number of countries that had at least one female head of state per region", x = "Year", y = "Number of countries at least one female head of state")
```

## wn_bothhouses_c

wn_bothhouses_c	measure	http://gapm.io/dwparl	v2	December 15 2020	Pecentage of women in parliaments	Pecentage of women in parliaments	Pecentage of women in parliaments	Percentage of national parliamentary seats held by women. Lower and upper houses combined	percent	governance

```{r countries-parliament-women-prc}
data$countries %>%
	select(region7eu, Code, Year, wn_bothhouses_c) %>%
	filter(!is.na(wn_bothhouses_c) & Year <= 2020 & !is.na(region7eu)) %>%
	rename(Value = wn_bothhouses_c) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7eu)) +
	geom_line(alpha = 0.45,  size = 0.85)
```


```{r}
data$countries %>%
	select(region7eu, Year, Code, female_hos, wn_bothhouses_c) %>%
	filter(!is.na(female_hos) & !is.na(wn_bothhouses_c)) %>%
	ggplot(aes(female_hos, wn_bothhouses_c)) +
	geom_hex()
```
