---
title: 'Dataset analysis: manual'
author: "Dereck de Mézquita"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), "./reports/", "/dataset-analysis-merged/dataset-analysis-merged.html")) })
output:
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
    # code_folding: hide
params:
  rmd: dataset-analysis-merged.Rmd
editor_options: 
  chunk_output_type: inline
---

<a download="dataset-analysis-merged.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 500, fig.width = 8, fig.height = 8, fig.align = "center")
```

# Libraries

```{r libraries, warning=FALSE, message=FALSE}
library("tidyverse")
options(dplyr.summarise.inform = FALSE)
# theme_set(theme(legend.position = "none"))
library("countrycode")
library("scales")

library("ddplot")
```

```{r modules, warning=FALSE, message=FALSE}
source("./modules/helper-functions.R")
```

# ggplot theme

```{r set-ggplot-defaults}
update_geom_defaults("line", list(alpha = 0.65, size = 0.65))

theme_set(theme_dereck())
```

# Load data

```{r load-data}
data <- readRDS("./outputs/clean_data.Rds")
meta <- read.csv("./data/ddf--gapminder--fasttrack/meta-data/ddf--concepts.csv")
```

# Dataset completeness

```{r function-invert-na-df}
is.not.na <- function(x) {
	return(!is.na(x))
}
```

The datasets collected and cleaned are of the following dimensions (rows, columns):

1. `data$pop`: `r dim(data$pop)`
1. `data$countries`: `r dim(data$countries)`
1. `data$global`: `r dim(data$global)`
1. `data$world_4region`: `r dim(data$world_4region)`

## Table preview

```{r preview-data-manual}
datatable(rbind(head(data$merged, 50), tail(data$merged, 50)))
```

## Dataset completeness {.tabset}

```{r barplot-completion-1, fig.height=15, results="asis"}
cat("### ","Dataset completion: merged", "\n")
{data$merged %>% 
		select(-Year, -Country, -Code, -region7, -region23, -eu15, -region7geoEU) %>%
		is.not.na() %>%
		as.data.frame() %>%
		colSums() / 64026 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	scale_x_discrete(expand = expansion(mult = c(0.015, 0.015))) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable", caption = "By Dereck de Mezquita") +
	coord_flip()
cat("\n\n")
```

```{r barplot-completion-2, results="asis"}
cat("### ","Dataset completion: manual", "\n")
{data$manual %>% 
		select(-Year, -Country, -Code, -region7, -region23, -eu15, -region7geoEU) %>%
		is.not.na() %>%
		as.data.frame() %>%
		colSums() / 58695 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	scale_x_discrete(expand = expansion(mult = c(0.015, 0.015))) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable", caption = "By Dereck de Mezquita") +
	coord_flip()
cat("\n\n")
```

```{r barplot-completion-3, fig.height=15, results="asis"}
cat("### ","Dataset completion: countries", "\n")
{data$countries %>% 
		select(-country, -time) %>%
		is.not.na() %>%
		as.data.frame() %>%
		colSums() / 58369 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	scale_x_discrete(expand = expansion(mult = c(0.015, 0.015))) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable", caption = "By Dereck de Mezquita") +
	coord_flip()
cat("\n\n")
```

# Meta data

Here is the meta data associated to these variables.

```{r table-meta-preview}
largeTable(meta, targs = c(3, 7, 9, 12))
```

## pop_total

```{r pop-line-world}
data$merged %>%
	rename(Value = pop_total) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7geoEU)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 20, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5, alpha = 0.75)) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.75))) + # make alpha coloured legend darker
	labs(title = "World population per country", subtitle = "Coloured per region; red dotted line: projected", x = "Year", y = "Population", colour = "Region", caption = "By Dereck de Mezquita")

data$merged %>%
	rename(Value = pop_total) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), group = Code, colour = region7geoEU)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 20, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5, alpha = 0.75)) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.75))) + # make alpha coloured legend darker
	labs(title = "World population per country", subtitle = "Coloured per region; red dotted line: projected", x = "Year", y = "log2(Population)", colour = "Region", caption = "By Dereck de Mezquita")
```

```{r pop-line-world-facet}
data$merged %>%
	rename(Value = pop_total) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = Code)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.175, 0.175))) +
	scale_y_continuous(n.breaks = 20, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.75))) + # make alpha coloured legend darker
	labs(title = "World population per country", subtitle = "Faceted per region; red dotted line: projected", x = "Year", y = "Population", colour = "Region", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")

data$merged %>%
	rename(Value = pop_total) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), group = Code, colour = Code)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.175, 0.175))) +
	scale_y_continuous(n.breaks = 20, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.75))) + # make alpha coloured legend darker
	labs(title = "World population per country", subtitle = "Faceted per region; red dotted line: projected", x = "Year", y = "log2(Population)", colour = "Region", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")
```

```{r pop-line-per-regions-sums}
data$merged %>%
	rename(Value = pop_total) %>%
	filter(!is.na(Value)) %>%
	group_by(Year, region7geoEU) %>%
	summarise(Value = sum(Value)) %>%
	ggplot(aes(Year, Value, colour = region7geoEU)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_x_continuous(n.breaks = 17, expand = expansion(mult = c(0.45, 0.45))) +
	scale_y_continuous(n.breaks = 17, labels = function(x) {
		# paste0(substr(format(x, scientific = FALSE), 1, 2), " B")
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	directlabels::geom_dl(aes(label = region7geoEU), method = list(directlabels::dl.trans(x = x + 0.2), "last.qp", cex = 0.75)) +
	directlabels::geom_dl(aes(label = region7geoEU), method = list(directlabels::dl.trans(x = x - 0.2), "first.qp", cex = 0.75)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "World population aggregated per region", subtitle = "Coloured per region; red dotted line: projected", x = "Year", y = "Population", colour = "Region", caption = "By Dereck de Mezquita")

data$merged %>%
	rename(Value = pop_total) %>%
	filter(!is.na(Value)) %>%
	group_by(Year, region23) %>%
	summarise(Value = sum(Value)) %>%
	ggplot(aes(Year, Value, colour = region23)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_x_continuous(n.breaks = 17, expand = expansion(mult = c(0.45, 0.45))) +
	scale_y_continuous(n.breaks = 17, labels = function(x) {
		# paste0(substr(format(x, scientific = FALSE), 1, 2), " B")
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	directlabels::geom_dl(aes(label = region23), method = list(directlabels::dl.trans(x = x + 0.2), "last.qp", cex = 0.75)) +
	directlabels::geom_dl(aes(label = region23), method = list(directlabels::dl.trans(x = x - 0.2), "first.qp", cex = 0.75)) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "World population aggregated per region", subtitle = "Coloured per region; red dotted line: projected", x = "Year", y = "Population", colour = "Region", caption = "By Dereck de Mezquita")
```

```{r pop-line-per-total-sums}
data$merged %>%
	rename(Value = pop_total) %>%
	filter(!is.na(Value)) %>%
	group_by(Year) %>% 
	summarise(Value = sum(Value)) %>%
	mutate(colour = ifelse(Year <= 2020, "#444444", "red"),
			 type = ifelse(Year <= 2020, "solid", "dashed")) %>%
	ggplot(aes(Year, Value, colour = colour, linetype = type)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_color_identity() +
	scale_linetype_identity() +
	scale_x_continuous(n.breaks = 30) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Total world population; red dotted line: projected", x = "Year", y = "Population - Billions", caption = "By Dereck de Mezquita") 

data$merged %>%
	select(Year, pop_total) %>%
	rename(Population = pop_total) %>% 
	drop_na() %>%
	group_by(Year) %>% 
	summarise(Population = sum(Population)) %>%
	mutate(colour = ifelse(Year <= 2020, "#444444", "red"),
			 type = ifelse(Year <= 2020, "solid", "dashed")) %>%
	ggplot(aes(Year, log2(Population), colour = colour, linetype = type)) +
	geom_line() +
	geom_vline(xintercept = 2020, linetype = "dotted", colour = "red") +
	scale_color_identity() +
	scale_linetype_identity() +
	scale_x_continuous(n.breaks = 30) +
	scale_y_continuous(n.breaks = 30, labels = function(x) { # calculate log2 scale labels manually; set aes(log2(data))
		scales::label_number_si(accuracy = 0.1)(2 ^ x) # more n.breaks when using log2 scale
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Total world population; red dotted line: projected", x = "Year", y = "log2(Population - Billions)", caption = "By Dereck de Mezquita")
```


# gdp

PPP GDP is gross domestic product converted to international dollars using purchasing power parity rates. A PPP dollar has the same purchasing power over GDP as the U.S. dollar has in the United States. Data are in constant 2005 international dollars.

```{r gpd-per-country}
data$merged %>%
	rename(Value = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7geoEU)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 30, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.title = element_text(vjust = 0.5)) +
	labs(title = "GDP per country", subtitle = "Coloured by region; constant 2005 international dollars", x = "Year", y = "Total GDP", colour = "Region", caption = "By Dereck de Mezquita")

data$merged %>%
	rename(Value = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), group = Code, colour = region7geoEU)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 30, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.title = element_text(vjust = 0.5)) +
	labs(title = "GDP per country", subtitle = "Coloured by region; constant 2005 international dollars", x = "Year", y = "log2(Total GDP)", colour = "Region", caption = "By Dereck de Mezquita")
```

```{r gpd-per-country-facet}
data$merged %>%
	rename(Value = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.25, 0.25))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	labs(title = "GDP per country; constant 2005 international dollars", x = "Year", y = "log2(Total GDP)", colour = "Region", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")

data$merged %>%
	rename(Value = tot_gdp_inflation_adj_2005) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), group = Code, colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.25, 0.25))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	labs(title = "GDP per country; constant 2005 international dollars", x = "Year", y = "log2(Total GDP)", colour = "Region", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")
```

```{r gdp-per-capita}
data$merged %>%
	rename(Value = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7geoEU)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75)) +
	scale_x_continuous(n.breaks = 30, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.title = element_text(vjust = 0.5)) +
	labs(title = "GDP/income income per capita per country", subtitle = "Coloured by region; constant 2011 international dollars", x = "Year", y = "GDP/income per capita", colour = "Region", caption = "By Dereck de Mezquita")

data$merged %>%
	rename(Value = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), group = Code, colour = region7geoEU)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 30, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 30, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.title = element_text(vjust = 0.5)) +
	labs(title = "GDP/income income per capita per country", subtitle = "Coloured by region; constant 2011 international dollars", x = "Year", y = "log2(GDP/income per capita)", colour = "Region", caption = "By Dereck de Mezquita")
```


```{r gdp-per-capita-facet}
data$merged %>%
	rename(Value = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.265, 0.265))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	labs(title = "GDP/income income per capita per region; constant 2011 international dollars", x = "Year", y = "GDP/income per capita", colour = "Region", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")

data$merged %>%
	rename(Value = gdp_per_capita_inflation_adj_2011) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), group = Code, colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.75, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.75, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.265, 0.265))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	labs(title = "GDP/income income per capita per region; constant 2011 international dollars", x = "Year", y = "log2(GDP/income per capita)", colour = "Region", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")
```


## avg_life_expect_yrs

```{r avg-life}
data$merged %>%
	rename(Value = avg_life_expect_yrs) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7geoEU)) +
	geom_line() +
	geom_vline(xintercept = 2020, colour = "red", linetype = "dashed") + 
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 20) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank(), legend.title = element_text(vjust = 0.5)) +
	labs(title = "Average life expectancy per country", subtitle = "Coloured by region", x = "Year", y = "Average life expectancy", colour = "Region", caption = "By Dereck de Mezquita")
```

```{r avg-life-aggregate}
data$merged %>%
	rename(Value = avg_life_expect_yrs) %>%
	filter(!is.na(Value)) %>%
	group_by(Year, region7geoEU) %>%
	summarise(Value = mean(Value)) %>%
	ggplot(aes(Year, Value, group = region7geoEU, colour = region7geoEU)) +
	geom_line() +
	geom_vline(xintercept = 2020, colour = "red", linetype = "dashed") +
	directlabels::geom_dl(aes(label = region7geoEU), method = list(directlabels::dl.trans(x = x + 0.2), "last.qp", cex = 0.75)) +
	directlabels::geom_dl(aes(label = region7geoEU), method = list(directlabels::dl.trans(x = x - 0.2), "first.qp", cex = 0.75)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.475, 0.475))) +
	scale_y_continuous(n.breaks = 20) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank(), legend.title = element_text(vjust = 0.5)) +
	labs(title = "Average life expectancy per region", subtitle = "Coloured by region", x = "Year", y = "Average life expectancy", colour = "Region", caption = "By Dereck de Mezquita")
```



```{r avg-life-facet}
data$merged %>%
	rename(Value = avg_life_expect_yrs) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = Code)) +
	geom_line() +
	geom_vline(xintercept = 2020, colour = "red", linetype = "dashed") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.5, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.5, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 15, expand = expansion(mult = c(0.2, 0.2))) +
	scale_y_continuous(n.breaks = 15) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.95))) +
	labs(title = "Average life expectancy per country", subtitle = "Coloured by region", x = "Year", y = "Average life expectancy", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")
```

## gini

Gini shows income inequality in a society. A higher number means more inequality.

gini	measure	http://gapm.io/ddgini	v3	December 5 2018	Gini	Gini coefficient		Gini shows income inequality in a society. A higher number means more inequality.

```{r gini-inequality-line}
data$merged %>%
	rename(Value = gini) %>%
	filter(!is.na(Value) & Year <= 2020, !is.na(region7geoEU)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7geoEU)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65, alpha = 0.75)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65, alpha = 0.75)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 20) +
	guides(colour = guide_legend(override.aes = list(alpha = 0.95))) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank(), legend.title = element_text(vjust = 0.5)) +
	labs(title = "Gini per country (inequality index)", subtitle = "A higher number means more inequality", x = "Year", y = "Gini index value", colour = "Region", caption = "By Dereck de Mezquita")
```

```{r gini-inequality-line-aggregate}
data$merged %>%
	rename(Value = gini) %>%
	filter(!is.na(Value) & Year <= 2020, !is.na(region7geoEU)) %>%
	group_by(Year, region7geoEU) %>%
	summarise(Value = mean(Value), region7geoEU) %>%
	ggplot(aes(Year, Value, group = region7geoEU, colour = region7geoEU)) +
	geom_line() +
	directlabels::geom_dl(aes(label = region7geoEU), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65)) +
	directlabels::geom_dl(aes(label = region7geoEU), method = list(directlabels::dl.trans(x = x - 0.2), "first.qp", cex = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.375, 0.375))) +
	scale_y_continuous(n.breaks = 20) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
	labs(title = "Average gini per region (inequality index)", subtitle = "A higher number means more inequality", x = "Year", y = "Gini index value", caption = "By Dereck de Mezquita")
```

## u5pop

Total number of children, aged 0 to 4 years old

u5pop	measure	http://gapm.io/du5pop	v1	January 31 2020	Under-five population	Number of under five years children	Under-five population	Total number of children, aged 0 to 4 years old.	people	population

```{r countries-u5pop-line}
data$merged %>%
	select(region7geoEU, Code, Year, u5pop) %>%
	rename(Value = u5pop) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65, alpha = 0.65)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65, alpha = 0.65)) +
	scale_x_continuous(n.breaks = 15,  expand = expansion(mult = c(0.25, 0.25))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	labs(title = "Total number of children per region: 0 - 5 yrs", x = "Year", y = "Number", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")

data$merged %>%
	select(region7geoEU, Code, Year, u5pop) %>%
	rename(Value = u5pop) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, log2(Value), colour = Code)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65, alpha = 0.65)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65, alpha = 0.65)) +
	scale_x_continuous(n.breaks = 15,  expand = expansion(mult = c(0.25, 0.25))) +
	scale_y_continuous(n.breaks = 15, labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(2 ^ x)
	}) +
	theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), panel.grid.minor = element_blank()) +
	labs(title = "Total number of children per region: 0 - 5 yrs", x = "Year", y = "Number", caption = "By Dereck de Mezquita") +
	facet_wrap(~ region7geoEU, scales = "free")
```

### u5pop/population

```{r countries-u5pop-prc}
data$merged %>% 
	select(region7geoEU, Code, Year, u5pop, pop_total) %>%
	filter(!is.na(u5pop) & !is.na(pop_total)) %>%
	mutate(Value = ((u5pop / pop_total) * 100)) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7geoEU)) +
	geom_line(alpha = 0.4, size = 0.95) +
	geom_vline(xintercept = 2020, colour = "red", linetype = "dashed") +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65, alpha = 0.65)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65, alpha = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 20, labels = function(x){
		paste0(x, "%")
	}) + # Add percent sign
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) + # make alpha coloured legend darker
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.title = element_text(vjust = 0.5)) +
	labs(title = "Percentage to country's population of children 0 - 5 yrs", subtitle = "Coloured by region", x = "Year", y = "Percent of children under 5 yrs", colour = "Region", caption = "By Dereck de Mezquita")
```

## female_hos

```{r countries-1-fhos}
data$merged %>%
	select(region7geoEU, Code, Year, female_hos) %>%
	filter(!is.na(female_hos) & Year <= 2019, !is.na(region7geoEU)) %>%
	mutate(Value = ifelse(female_hos == "Never had a female head of state", 0, 1)) %>%
	ggplot(aes(Year, Value, group = region7geoEU, fill = region7geoEU)) +
	geom_bar(stat = "identity") +
	scale_y_continuous(n.breaks = 20) +
	scale_x_continuous(n.breaks = 20) +
	theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.title = element_text(vjust = 0.5)) +
	labs(title = "Number of countries at least one female head of state", subtitle = "Coloured per region", x = "Year", y = "Number of countries at least one female head of state", fill = "Region", caption = "By Dereck de Mezquita")
```

## wn_bothhouses_c

wn_bothhouses_c	measure	http://gapm.io/dwparl	v2	December 15 2020	Pecentage of women in parliaments	Pecentage of women in parliaments	Pecentage of women in parliaments	Percentage of national parliamentary seats held by women. Lower and upper houses combined	percent	governance

```{r countries-parliament-women-prc}
data$merged %>%
	rename(Value = wn_bothhouses_c) %>%
	filter(!is.na(Value) & Year <= 2020 & !is.na(region7geoEU)) %>%
	ggplot(aes(Year, Value, group = Code, colour = region7geoEU)) +
	geom_line() +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x + 0.2), "last.points", cex = 0.65, alpha = 0.65)) +
	directlabels::geom_dl(aes(label = Code), method = list(directlabels::dl.trans(x = x - 0.2), "first.points", cex = 0.65, alpha = 0.65)) +
	scale_x_continuous(n.breaks = 20, expand = expansion(mult = c(0.1, 0.1))) +
	scale_y_continuous(n.breaks = 20, labels = function(x) {
		paste0(x, "%")
	}) +
	guides(color = guide_legend(override.aes = list(alpha = 0.95))) + # make alpha coloured legend darker
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.title = element_text(vjust = 0.5)) +
	labs(title = "Percentage of women to parliaments per country", subtitle = "Coloured by region", x = "Year", y = "Percent of women in parliaments", colour = "Region", caption = "By Dereck de Mezquita")
```

## polrights_fh

polrights_fh measure http://gapm. io/dfreedom_fh v1 April 24 2019 Political Rights index (FH) Political Rights index (FH) "Political rights is assessed by 10 indicators asked in form of questions and grouped into 3 subcategories

```{r}
data$merged %>%
	rename(Value = polrights_fh) %>%
	filter(!is.na(Value)) %>%
	ggplot(aes(Year, Value, group = region7geoEU, fill = region7geoEU)) +
	geom_bar(stat = "identity") +
	scale_x_continuous(n.breaks = 20) +
	scale_y_continuous(n.breaks = 20) +
	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

