---
title: 'Dataset choropleths: merged'
author: "Dereck de Mézquita"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), "./reports/", "/dataset-choropleths-merged/dataset-choropleths-merged.html")) })
output:
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
    # code_folding: hide
params:
  rmd: dataset-choropleths-merged.Rmd
editor_options: 
  chunk_output_type: inline
---

<a download="dataset-choropleths-merged.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 500, fig.width = 8, fig.height = 8, fig.align = "center")
```

# Libraries

```{r libraries, warning=FALSE, message=FALSE}
library("tidyverse")
library("ddplot")

library("geojsonio")
library("broom")

library("viridis")
```

# ggplot theme

```{r set-ggplot-defaults}
update_geom_defaults("line", list(alpha = 0.65, size = 0.65))

theme_set(theme_dereck())
```

# Load data

```{r load-data}
data <- readRDS("./outputs/clean_data.Rds")$merged
```

## Print data

```{r preview-data}
data
```

# Download the .geoJSON

```{bash download-geo-json}
cd data/mapping-data
wget https://raw.githubusercontent.com/deldersveld/topojson/master/world-countries.json
wget https://raw.githubusercontent.com/johan/world.geo.json/master/countries.geo.json
```

## Load and preview geoJSON

```{r read-geojson}
# plot(geojsonio::geojson_read("./data/mapping-data/world-countries.json", what = "sp"))
spdf <- geojsonio::geojson_read("./data/mapping-data/countries.geo.json", what = "sp")

# plot(spdf)
```

## Convert to tidy geoJSON

```{r tidy-geojson, warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
spdf_tdy <- broom::tidy(spdf, region = "id")

spdf_tdy %>%
	ggplot() +
	geom_polygon(aes(x = long, y = lat, group = group), fill = "white", colour = "gray") +
	theme_void() +
	coord_map()
```

# Preview data variation

```{r func-is-not-na}
is.not.na <- function(x) {
	return(!is.na(x))
}
```

```{r completeness, fig.height=15}
{data %>% 
		filter(Year == 2015) %>%
		select(-Year, -Country, -Code, -region7geoEU, -region7, -region7eu, -region23, -eu15) %>% 
		is.not.na() %>%
		as.data.frame() %>%
		colSums() / 214 * 100} %>%
	as.list() %>%
	data.frame() %>%
	t() %>%
	as.data.frame() %>%
	rownames_to_column() %>%
	ggplot(aes(reorder(rowname, V1), V1, fill = V1)) +
	geom_bar(stat = "identity") +
	scale_fill_gradient(low = "blue", high = "red", limits = c(0, 100), breaks = c(0, 25, 50, 75, 100))  +
	geom_text(aes(label = glue::glue("{round(V1)}%")), hjust = -0.35, colour = "black") + # , angle = 90
	scale_y_continuous(expand = expansion(mult = c(0, .2)), n.breaks = 15) +
	scale_x_discrete(expand = expansion(mult = c(0.015, 0.015))) +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
	labs(title = "Completeness of variables; year 2015", subtitle = "Percent of non NA values per variable", x = "Percentage completeness", y = "Variable", caption = "By Dereck de Mezquita") +
	coord_flip()
```


```{r data-cols}
colnames(data)
```

ilevels4_wb	string	http://gapm.io/dilevels_wb	v2	September 17 2019	4 income levels	4 income levels (World Bank)		Four country income levels classified by the World Bank based on GNI per capita in US$ (Atlas methodology)

```{r bar-pop-chng-country, fig.height=25, fig.width=10}
data %>%
	select(Year, Country, region7geoEU, pop_total) %>%
	filter(!is.na(pop_total) & Year %in% c(2010, 2020)) %>%
	group_by(Country) %>%
	arrange(Year, .by_group = TRUE) %>%
	mutate(pop_chng = (pop_total / lag(pop_total) - 1) * 100) %>%
	filter(!is.na(pop_chng)) %>%
	ggplot(aes(reorder(Country, pop_chng), pop_chng, fill = region7geoEU)) +
	geom_bar(stat = "identity") +
	scale_y_continuous(n.breaks = 20, labels = function(x){
		paste0(x, "%")
	}) +
	coord_flip() +
	theme(panel.grid.minor = element_blank(), legend.title = element_text(vjust = 0.5)) +
	labs(title = "Population change per country 2010 - 2020", x = "Percent change 2010 - 2020", y = "Country", fill = "Region")
```

```{r 2015-data-merge-map}
map_data <- data %>%
	select(Year, Code, Country, region7geoEU, pop_total) %>%
	filter(!is.na(pop_total) & Year %in% c(2010, 2020)) %>%
	group_by(Country) %>%
	arrange(Year, .by_group = TRUE) %>%
	mutate(prc_chng = (pop_total / lag(pop_total) - 1) * 100,
			 tot_chng = (pop_total - lag(pop_total))) %>%
	filter(!is.na(prc_chng) | !is.na(tot_chng)) %>%
	full_join(spdf_tdy, by = c("Code" = "id"))


map_data$prc_chng[is.na(map_data$prc_chng)] <- 0.001
map_data$tot_chng[is.na(map_data$tot_chng)] <- 0.001
map_data$tot_chng[map_data$tot_chng == 0] <- 0.001
```

```{r bar-prc_chng-dist}
map_data %>%
	filter(!is.na(long) | !is.na(lat)) %>%
	select(Code, prc_chng) %>%
	distinct() %>%
	ggplot(aes(reorder(Country, prc_chng), prc_chng)) +
	geom_bar(stat = "identity") +
	scale_y_continuous(n.breaks = 15) +
	coord_flip()
```

```{r check-min-max}
min(map_data$prc_chng)
max(map_data$prc_chng)
```

```{r chor-1-prc-chng, fig.height=8, fig.width=10}
map_data %>%
	ggplot() +
	geom_polygon(aes(x = long, y = lat, group = group, fill = prc_chng)) +
	theme_void() +
	coord_map() +
	# scale_y_continuous(n.breaks = 40) +
	scale_fill_gradient(limits = c(-20, 70), breaks = c(-20, -10, 0, 10, 20, 30, 40, 50, 60, 70, 80), labels = function(x) {
		paste0(x, "%")
	}) + # low = "blue", high = "yellow", 
	labs(title = "Percent change in population 2010 - 2020", caption = "By: Dereck de Mezquita", fill = "Change") +
	theme(legend.position = "right")
```

```{r chor-2-prc-chng, fig.height=8, fig.width=10}
map_data %>%
	ggplot() +
	geom_polygon(aes(fill = prc_chng, x = long, y = lat, group = group), size = 0, alpha = 0.9) +
	theme_void() +
	coord_map() +
	scale_fill_viridis(n.breaks = 10, guide = guide_legend(keyheight = unit(3, units = "mm"), keywidth = unit(5, units = "mm"), nrow = 5), labels = function(x) {
		paste0(x, "%")
	}) + # label.position = "bottom", title.position = "top"
	labs(title = "Percent change in population 2010 - 2020", caption = "By: Dereck de Mezquita", fill = element_blank()) +
	theme(plot.title = element_text(size = 15, hjust = 0.01, margin = margin(r = 10, l = 2, b = 0.5, unit = "cm")), plot.caption = element_text(margin = margin(15, 50, unit = "mm")), legend.position = c(0.5, -0.075))
```

```{r chor-3-prc-chng, fig.height=8, fig.width=10}
map_data %>%
	ggplot() +
	geom_polygon(aes(fill = tot_chng, x = long, y = lat, group = group), size = 0, alpha = 0.9) +
	theme_void() +
	coord_map() +
	scale_fill_viridis(n.breaks = 10, guide = guide_legend(keyheight = unit(3, units = "mm"), keywidth = unit(5, units = "mm"), nrow = 5), labels = function(x) {
		scales::label_number_si(accuracy = 0.1)(x)
		# scales::label_number_si(accuracy = 0.1)(x)
	}) + # label.position = "bottom", title.position = "top"
	labs(title = "Total change in population 2010 - 2020", caption = "By: Dereck de Mezquita", fill = element_blank()) +
	theme(plot.title = element_text(size = 15, hjust = 0.01, margin = margin(r = 10, l = 2, b = 0.5, unit = "cm")), plot.caption = element_text(margin = margin(15, 50, unit = "mm")), legend.position = c(0.5, -0.075))
```

