---
title: "USA: dataset preprocessing"
author: "Dereck de Mézquita"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), "./reports/", "us-state-county-dataset-preprocessing/us-state-county-dataset-preprocessing.html")) })
output:
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
params:
  rmd: us-state-county-dataset-preprocessing.Rmd
editor_options: 
  chunk_output_type: inline
---

<a download="us-state-county-dataset-preprocessing.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

# Libraries

```{r libraries, warning=FALSE, message=FALSE}
library("tidyverse")
# library("readr")
# library("openxlsx")
library("readxl")
library("vroom")

# library("httr")

library("tidycensus")
```

```{r load-modules, warning=FALSE, message=FALSE}
source("./modules/helper-functions.R")
```

# Description

The goal of this document is to collect data from the census bureau and others relating to the population of the United States of America. These data should cover a variety of attributes such as GDP and education, at the county level preferably.

This document is for collecting and processing said data. Other documents will be created for analysing this data.

# FIP codes per county

```{r}
data("fips_codes")

fips_codes <- fips_codes %>%
   mutate(fips = as.character(paste0(state_code, county_code)), county = trimws(gsub(" County", "", county)))

names(fips_codes) <- c("abbrv", "stcode", "state", "ctycode", "county", "fips")
```

```{r}
datatable(fips_codes)
```

```{r}
write.csv(fips_codes[, c("fips", "abbrv", "state", "county")], "./outputs/united-states-of-america/fip_codes.csv", row.names = FALSE)
```

# GDP per county: 2001 - 2019

"CAGDP2 Gross domestic product (GDP) by county and metropolitan area 1/"
"Gross Domestic Product (GDP): All industry total (Thousands of dollars)" 
"Bureau of Economic Analysis" 
"County"

"Legend / Footnotes:"
"1/ Gross Domestic Product (GDP) is in thousands of current dollars (not adjusted for inflation). Industry detail is based on the 2012 North American Industry Classification System (NAICS)."
"* Broomfield County, CO, was created from parts of Adams, Boulder, Jefferson, and Weld counties effective November 15, 2001. Estimates for Broomfield county begin with 2002."
"* Estimates from 2008 forward separate Skagway-Hoonah-Angoon Census Area into Skagway Municipality and Hoonah-Angoon Census Area. Estimates from 2009 forward separate Wrangell-Petersburg Census Area into Petersburg Census Area and Wrangell City and Borough. In addition, a part of the Prince of Wales-Outer Ketchikan Census Area was annexed by Ketchikan Gateway Borough and part (Meyers Chuck Area) was included in the new Wrangell City and Borough. The remainder of the Prince of Wales-Outer Ketchikan Census Area was renamed Prince of Wales-Hyder Census Area. Petersburg Borough was created from part of former Petersburg Census Area and part of Hoonah-Angoon Census Area for 2013 forward. Prince of Wales-Hyder Census Area added part of the former Petersburg Census Area beginning in 2013. For years 2009-2012, Petersburg Borough reflects the geographic boundaries of the former Petersburg Census Area. Wade Hampton Census Area was renamed Kusilvak Census Area on July 1, 2015."
"* Virginia combination areas consist of one or two independent cities with 1980 populations of less than 100,000 combined with an adjacent county. The county name appears first, followed by the city name(s). Separate estimates for the jurisdictions making up the combination area are not available. Bedford County, VA includes the independent city of Bedford for all years."
"* Shannon County, SD was renamed to Oglala Lakota County, SD on May 1, 2015."
"* Kalawao County, Hawaii is combined with Maui County. Separate estimates for the jurisdictions making up the combination areas are not available."
"Metropolitan Areas are defined (geographically delineated) by the Office of Management and Budget (OMB) bulletin no. 20-01 issued March 6, 2020."
"(NA) Not available."
"Last updated: December 9, 2020 -- new statistics for 2019, revised statistics for 2001-2018."

Source: https://apps.bea.gov/iTable/iTable.cfm?reqid=70&step=1&acrdn=5

```{r}
gdp <- lapply(list.files("./data/united-states-of-america/per-county-gdp-bea", pattern = "\\.csv$", full.names = TRUE), read.csv)

gdp <- lapply(gdp, function(x) {
   names(x)[1] <- "fips"
   x[,2] <- NULL
   return(x)
})

gdp <- Reduce(function(...) {
   merge(..., all.x = FALSE, by = "fips")
}, gdp); # gdp
```

```{r}
dim(gdp); datatable(head(gdp, 30))
```

```{r}
write.csv(gdp, "./outputs/united-states-of-america/gdp_2001_2019.csv", row.names = FALSE)
```

# Census data

https://www.census.gov/data/tables/time-series/demo/popest/2010s-counties-total.html#par_textimage

Population, Population Change, and Estimated Components of Population Change: April 1, 2010 to July 1, 2019 (CO-EST2019-alldata):
   - https://www.census.gov/data/tables/time-series/demo/popest/2010s-counties-total.html
File layout:
   - https://www2.census.gov/programs-surveys/popest/technical-documentation/file-layouts/2010-2019/co-est2019-alldata.pdf

```{r}
if(!file.exists("./data/united-states-of-america/per-county-pops-census/census_2019.csv")) {
   census <- read.csv("https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv", colClasses = c(STATE = "character", COUNTY = "character")); # census
   write.csv(census, "./data/united-states-of-america/per-county-pops-census/census_2019.csv", row.names = FALSE)
} else {
   census <- read.csv("./data/united-states-of-america/per-county-pops-census/census_2019.csv", row.names = 1, colClasses = c(STATE = "character", COUNTY = "character")); # census
}
```

## Concatenate state code to county code

This to create the full FIP codes.

```{r}
census <- census %>%
   mutate(fips = paste0(STATE, COUNTY))

census <- data.frame(FIPS = census$fips, census[,1:(ncol(census)-1)])
```

REGION

1 = Northeast
2 = Midwest 
3 = South
4 = West

DIVISION

1 = New England
2 = Middle Atlantic
3 = East North Central
4 = West North Central 
5 = South Atlantic
6 = East South Central 
7 = West South Central 
8 = Mountain
9 = Pacific

```{r}
census <- census %>%
   mutate(REGION = case_when(REGION == 1 ~ "Northeast", 
                             REGION == 2 ~ "Midwest", 
                             REGION == 3 ~ "South", 
                             REGION == 4 ~ "West"),
          DIVISION = case_when(DIVISION == 1 ~ "New England", 
                             DIVISION == 2 ~ "Middle Atlantic", 
                             DIVISION == 3 ~ "East North Central", 
                             DIVISION == 4 ~ "West North Central",
                             DIVISION == 5 ~ "South Atlantic",
                             DIVISION == 6 ~ "East South Central",
                             DIVISION == 7 ~ "West South Central",
                             DIVISION == 8 ~ "Mountain",
                             DIVISION == 9 ~ "Pacific")); # census
```

## Remove state rows

```{r}
# census[census$FIPS %in% fips_codes$fips,]

census <- census %>%
   filter(!as.numeric(COUNTY) == 0)
```


## Join to FIPs

```{r}
sum(fips_codes$fips %in% census$FIPS)

census <- census %>% 
   select(-STNAME, -CTYNAME, -SUMLEV, -STATE, -COUNTY)
```

## Subdividing census data

```{r}
census_data <- list()

census_data$population <- census %>%
   select(FIPS, REGION, DIVISION, CENSUS2010POP, POPESTIMATE2010, POPESTIMATE2011, POPESTIMATE2012, POPESTIMATE2013, POPESTIMATE2014, POPESTIMATE2015, POPESTIMATE2016, POPESTIMATE2017, POPESTIMATE2018, POPESTIMATE2019)

census_data$births <- census %>%
   select(FIPS, REGION, DIVISION, BIRTHS2010, BIRTHS2011, BIRTHS2012, BIRTHS2013, BIRTHS2014, BIRTHS2015, BIRTHS2016, BIRTHS2017, BIRTHS2018, BIRTHS2019)

census_data$deaths <- census %>%
   select(FIPS, REGION, DIVISION, DEATHS2010, DEATHS2011, DEATHS2012, DEATHS2013, DEATHS2014, DEATHS2015, DEATHS2016, DEATHS2017, DEATHS2018, DEATHS2019)

census_data$int_migration <- census %>%
   select(FIPS, REGION, DIVISION, INTERNATIONALMIG2010, INTERNATIONALMIG2011, INTERNATIONALMIG2012, INTERNATIONALMIG2013, INTERNATIONALMIG2014, INTERNATIONALMIG2015, INTERNATIONALMIG2016, INTERNATIONALMIG2017, INTERNATIONALMIG2018, INTERNATIONALMIG2019)

census_data$dom_migration <- census %>%
   select(FIPS, REGION, DIVISION, DOMESTICMIG2010, DOMESTICMIG2011, DOMESTICMIG2012, DOMESTICMIG2013, DOMESTICMIG2014, DOMESTICMIG2015, DOMESTICMIG2016, DOMESTICMIG2017, DOMESTICMIG2018, DOMESTICMIG2019)

census_data$net_migration <- census %>%
   select(FIPS, REGION, DIVISION, NETMIG2010, NETMIG2011, NETMIG2012, NETMIG2013, NETMIG2014, NETMIG2015, NETMIG2016, NETMIG2017, NETMIG2018, NETMIG2019)
```

```{r}
list_df_columns_lower <- function(data) {
   data <- lapply(data, function(x) {
      colnames(x) <- tolower(colnames(x))
      return(x)
   })
   return(data)
}
```

```{r}
census_data <- list_df_columns_lower(census_data); # census_data
```

```{r}
sapply(names(census_data), function(x) {
   write.csv(census_data[[x]], glue::glue("./outputs/united-states-of-america/census_{x}.csv"), row.names = FALSE)
})
```


# Economic Research Service United States Department of Agriculture

https://www.ers.usda.gov/data-products/county-level-data-sets/download-data/

- Unemployment.xls
- Education.xls
- PovertyEstimates.xls

```{r}
usda <- lapply(list.files("./data/united-states-of-america/per-county-econ-usda/", full.names = TRUE), read_excel, sheet = 1)
names(usda) <- gsub(".xls", "", list.files("./data/united-states-of-america/per-county-econ-usda/"), perl = TRUE)

usda <- lapply(usda, as.data.frame)
usda <- list_df_columns_lower(usda)
usda <- lapply(usda, function(x) {
      rename_with(x, .fn = function(x) {
         tolower(gsub("'|,|\\(|\\)", "", gsub(" |-", "_", x)))
      })
}); # usda
```

```{r}
usda <- lapply(usda, function(x) {
   x <- x[-1,]
})
```

```{r}
names(usda$Education)[1] <- "fipstxt"
names(usda$Education)[2] <- "stabr"
```

```{r}
usda <- lapply(usda, function(x) {
	data.frame(fip_st = (substr(x$fipstxt, 1, 2)), fip_cty = (substr(x$fipstxt, 3, length(x$fipstxt))), x)
})
```

```{r}
usda <- lapply(usda, function(x) {
	x %>% 
		filter(!(as.numeric(fip_cty) == 0)) %>%
		select(-fip_st, -fip_cty)
})
```

## Education

```{r}
usda$Education <- usda$Education %>%
   select(fipstxt,
          less_than_a_high_school_diploma_1970, high_school_diploma_only_1970, some_college_1_3_years_1970, four_years_of_college_or_higher_1970,
          less_than_a_high_school_diploma_1980, high_school_diploma_only_1980, some_college_1_3_years_1980, four_years_of_college_or_higher_1980,
          less_than_a_high_school_diploma_1990, high_school_diploma_only_1990, some_college_or_associates_degree_1990, bachelors_degree_or_higher_1990,
          less_than_a_high_school_diploma_2000, high_school_diploma_only_2000, some_college_or_associates_degree_2000, bachelors_degree_or_higher_2000,
          less_than_a_high_school_diploma_2014_18, high_school_diploma_only_2014_18, some_college_or_associates_degree_2014_18, bachelors_degree_or_higher_2014_18,
          percent_of_adults_with_less_than_a_high_school_diploma_1970, percent_of_adults_with_a_high_school_diploma_only_1970, percent_of_adults_completing_some_college_1_3_years_1970, percent_of_adults_completing_four_years_of_college_or_higher_1970,
          percent_of_adults_with_less_than_a_high_school_diploma_1980, percent_of_adults_with_a_high_school_diploma_only_1980, percent_of_adults_completing_some_college_1_3_years_1980, percent_of_adults_completing_four_years_of_college_or_higher_1980,
          percent_of_adults_with_less_than_a_high_school_diploma_1990, percent_of_adults_with_a_high_school_diploma_only_1990, percent_of_adults_completing_some_college_or_associates_degree_1990, percent_of_adults_with_a_bachelors_degree_or_higher_1990,
          percent_of_adults_with_less_than_a_high_school_diploma_2000, percent_of_adults_with_a_high_school_diploma_only_2000, percent_of_adults_completing_some_college_or_associates_degree_2000, percent_of_adults_with_a_bachelors_degree_or_higher_2000,
          percent_of_adults_with_less_than_a_high_school_diploma_2014_18, percent_of_adults_with_a_high_school_diploma_only_2014_18, percent_of_adults_completing_some_college_or_associates_degree_2014_18, percent_of_adults_with_a_bachelors_degree_or_higher_2014_18)
```

```{r}
colnames(usda$Education) <- c("fips",
                              "no_high_1970", "high_only_1970", "some_college_1_3_yrs_1970", "four_years_college_or_more_1970",
                              "no_high_1980", "high_only_1980", "some_college_1_3_yrs_1980", "four_years_college_or_more_1980",
                              "no_high_1990", "high_only_1990", "some_college_associates_1990", "bachelors_degree_or_more_1990",
                              "no_high_2000", "high_only_2000", "some_college_associates_2000", "bachelors_degree_or_more_2000",
                              "no_high_2014_18", "high_only_2014_18", "some_college_associates_2014_18", "bachelors_degree_or_more_2014_18",
                              "pct_adults_no_high_1970", "pct_adults_only_high_1970", "pct_adults_some_college_1_3_yrs_1970", "pct_adults_four_yrs_college_or_more_1970",
                              "pct_adults_no_high_1980", "pct_adults_only_high_1980", "pct_adults_some_college_1_3_yrs_1980", "pct_adults_four_yrs_college_or_more_1980",
                              "pct_adults_no_high_1990", "pct_adults_only_high_1990", "pct_adults_some_college_associates_1990", "pct_adults_bachelors_or_more_1990",
                              "pct_adults_no_high_2000", "pct_adults_only_high_2000", "pct_adults_some_college_associates_2000", "pct_adults_bachelors_or_more_2000",
                              "pct_adults_no_high_2014_18", "pct_adults_only_high_2014_18", "pct_adults_some_college_associates_2014_18", "pct_adults_bachelors_or_more_2014_18")
```

```{r}
dim(usda$Education); datatable(head(usda$Education, 50))
```

## Poverty

```{r}
usda$PovertyEstimates <- usda$PovertyEstimates %>%
   select(fipstxt, povall_2018, pctpovall_2018)

colnames(usda$PovertyEstimates) <- c("fips", "tot_indiv_poverty_2018", "pct_povall_2018"); # usda$PovertyEstimates
```

```{r}
dim(usda$PovertyEstimates); datatable(head(usda$PovertyEstimates))
```

## Employment

```{r}
usda$Unemployment <- usda$Unemployment %>%
   select(fipstxt,
          civilian_labor_force_2000, employed_2000, unemployed_2000,
          civilian_labor_force_2001, employed_2001, unemployed_2001,
          civilian_labor_force_2002, employed_2002, unemployed_2002,
          civilian_labor_force_2003, employed_2003, unemployed_2003,
          civilian_labor_force_2004, employed_2004, unemployed_2004,
          civilian_labor_force_2005, employed_2005, unemployed_2005,
          civilian_labor_force_2006, employed_2006, unemployed_2006,
          civilian_labor_force_2007, employed_2007, unemployed_2007,
          civilian_labor_force_2008, employed_2008, unemployed_2008,
          civilian_labor_force_2009, employed_2009, unemployed_2009,
          civilian_labor_force_2010, employed_2010, unemployed_2010,
          civilian_labor_force_2011, employed_2011, unemployed_2011,
          civilian_labor_force_2012, employed_2012, unemployed_2012,
          civilian_labor_force_2013, employed_2013, unemployed_2013,
          civilian_labor_force_2014, employed_2014, unemployed_2014,
          civilian_labor_force_2015, employed_2015, unemployed_2015,
          civilian_labor_force_2016, employed_2016, unemployed_2016,
          civilian_labor_force_2017, employed_2017, unemployed_2017,
          civilian_labor_force_2018, employed_2018, unemployed_2018,
          civilian_labor_force_2019, employed_2019, unemployed_2019,
          median_household_income_2018, med_hh_income_percent_of_state_total_2018,
          unemployment_rate_2000, unemployment_rate_2001, unemployment_rate_2002, 
          unemployment_rate_2003, unemployment_rate_2004, unemployment_rate_2005, 
          unemployment_rate_2006, unemployment_rate_2007, unemployment_rate_2008, 
          unemployment_rate_2009, unemployment_rate_2010, unemployment_rate_2011, 
          unemployment_rate_2012, unemployment_rate_2013, unemployment_rate_2014, 
          unemployment_rate_2015, unemployment_rate_2016, unemployment_rate_2017, 
          unemployment_rate_2018, unemployment_rate_2019) %>%
   rename(fips = fipstxt); # usda$Unemployment
```


```{r}
dim(usda$Unemployment); datatable(head(usda$Unemployment))
```

```{r}
sapply(names(usda), function(x) {
   write.csv(usda[[x]], glue::glue("./outputs/united-states-of-america/jobs_{tolower(x)}.csv"), row.names = FALSE)
})
```

# Merge all

```{r}
census_merge <- Reduce(function(...) {
   merge(..., all.x = FALSE) # , by = c("fips", "region", "division", "stname", "ctyname")
}, census_data); # census_merge

usda_merge <- Reduce(function(...) {
   merge(..., all.x = FALSE)
}, usda); # usda_merge

merged <- Reduce(function(...) {
   merge(..., all.x = FALSE)
}, list(fips_codes[, c("fips", "abbrv", "state", "county")], census_merge, gdp, usda_merge)); # merge
```

```{r}
write.csv(merged, file = "./outputs/united-states-of-america/all_merged_data.csv", row.names = FALSE)
```

```{r}
saveRDS(merged, "./outputs/united-states-of-america/all_merged_data.Rds")
```

## Race/ethnic origin

https://www.census.gov/data/tables/time-series/demo/popest/2010s-counties-detail.html

This is a very large dataset and requires some special attention.



```{bash download-geo-json}
if [ ! -f data/united-states-of-america/per-county-race-census/cc-est2019-alldata.Rds ]; then
   echo "File not found: downloading file."
   cd data/united-states-of-america/per-county-race-census/
	wget https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/asrh/cc-est2019-alldata.csv
else
   echo "File already exists: not downloading"
fi
````

```{r}
if(!file.exists("./data/united-states-of-america/per-county-race-census/cc-est2019-alldata.Rds")) {
   race <- read.csv("./data/united-states-of-america/per-county-race-census/cc-est2019-alldata.csv", colClasses = c(STATE = "character", COUNTY = "character"))
   saveRDS(race, "./data/united-states-of-america/per-county-race-census/cc-est2019-alldata.Rds")
} else {
   race <- readRDS("./data/united-states-of-america/per-county-race-census/cc-est2019-alldata.Rds"); # census
}
```

```{bash}
if [ -f data/united-states-of-america/per-county-race-census/cc-est2019-alldata.Rds ]; then
   echo "RDS exists, removing CSV."
   rm data/united-states-of-america/per-county-race-census/cc-est2019-alldata.csv
else
   echo "Error RDS not found."
fi
```

```{r}
race <- race %>%
   mutate(fips = paste0(STATE, COUNTY))

race <- data.frame(FIPS = race$fips, race[,1:(ncol(race)-1)])
```

```{r}
# census[census$FIPS %in% fips_codes$fips,]

race <- race %>%
   filter(!as.numeric(COUNTY) == 0)
```

The key for the YEAR variable is as follows: 

- 1 = 4/1/2010 Census population
- 2 = 4/1/2010 population estimates base 
- 3 = 7/1/2010 population estimate
- 4 = 7/1/2011 population estimate 
- 5 = 7/1/2012 population estimate 
- 6 = 7/1/2013 population estimate 
- 7 = 7/1/2014 population estimate 
- 8 = 7/1/2015 population estimate 
- 9 = 7/1/2016 population estimate
- 10 = 7/1/2017 population estimate 
- 11 = 7/1/2018 population estimate 
- 12 = 7/1/2019 population estimate

The key for AGEGRP is as follows: 

- 0 = Total
- 1 = Age 0 to 4 years
- 2 = Age 5 to 9 years
- 3 = Age 10 to 14 years
- 4 = Age 15 to 19 years
- 5 = Age 20 to 24 years
- 6 = Age 25 to 29 years
- 7 = Age 30 to 34 years
- 8 = Age 35 to 39 years
- 9 = Age 40 to 44 years
- 10 = Age 45 to 49 years 
- 11 = Age 50 to 54 years 
- 12 = Age 55 to 59 years 
- 13 = Age 60 to 64 years 
- 14 = Age 65 to 69 years 
- 15 = Age 70 to 74 years 
- 16 = Age 75 to 79 years 
- 17 = Age 80 to 84 years 
- 18 = Age 85 years or older

```{r}
race <- race %>%
	filter(!YEAR %in% c(1, 2) & !AGEGRP == 0) %>%
	mutate(YEAR = case_when(YEAR == 3 ~ 2010,
									YEAR == 4 ~ 2011,
									YEAR == 5 ~ 2012,
									YEAR == 6 ~ 2013,
									YEAR == 7 ~ 2014,
									YEAR == 8 ~ 2015,
									YEAR == 9 ~ 2016,
									YEAR == 10 ~ 2017,
									YEAR == 11 ~ 2018,
									YEAR == 12 ~ 2019),
			 AGEGRP = case_when(AGEGRP == 1 ~ "0 - 4",
									 AGEGRP == 2 ~ "5 - 9",
									 AGEGRP == 3 ~ "10 - 14",
									 AGEGRP == 4 ~ "15 - 19",
									 AGEGRP == 5 ~ "20 - 24",
									 AGEGRP == 6 ~ "25 - 29",
									 AGEGRP == 7 ~ "30 - 34",
									 AGEGRP == 8 ~ "35 - 39",
									 AGEGRP == 9 ~ "40 - 44",
									 AGEGRP == 10 ~ "45 - 49", 
									 AGEGRP == 11 ~ "50 - 54",
									 AGEGRP == 12 ~ "55 - 59",
									 AGEGRP == 13 ~ "60 - 64",
									 AGEGRP == 14 ~ "65 - 69",
									 AGEGRP == 15 ~ "70 - 74",
									 AGEGRP == 16 ~ "75 - 79",
									 AGEGRP == 17 ~ "80 - 84",
									 AGEGRP == 18 ~ "85+"))
```

```{r}
race <- as.data.frame(list_df_columns_lower(list(race)))
```


```{r}
pvtd_race <- race %>%
	select(-sumlev, -state, -county, -stname, -ctyname) %>%
	pivot_longer(cols = tot_pop:hnac_female, names_to = "races")
```

```{r}
pvtd_race %>%
	filter(races %in% c("tot_male", "tot_female")) %>%
	group_by(fips, year, races) %>%
	summarise(total = sum(value))
```


```{r}
pvtd_race %>%
	filter(races %in% c("tot_male", "tot_female")) %>%
	group_by(fips, year, races) %>%
	summarise(total = sum(value)) %>%
	ggplot(aes(year, total, group = fips, colour = fips)) + 
	geom_bar(stat = "identity") + 
	theme(legend.position = "none")
```



# Tidy up data

```{r}

```


# Session info

```{r}
sessionInfo()
```


